Approximation as Differentiation

### 2025-04-15 appendix/notes/biproduct.tex
Assume universal quantification implicit.
Identities in LatGal.
mathcal for metavariables that range over categories.

\begin{definition}[Biproduct]
Suppose $C$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is an
object $A \biprod B$ of $C$ together with morphisms

@@

We say a $\CMon$-enriched $C$ is \emph{semi-additive} if it also has finite biproducts; having products or
coproducts is sufficient.

\begin{proposition}
\label{prop:biproduct:from-product-or-coproduct}
Suppose $C$ is $\CMon$-enriched.
\begin{itemize}
\item If $C$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
\item If $C$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
\prodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $C$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $C$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

\begin{proposition}
\label{prop:biproduct:prod-coprod}
Suppose $C$ a category with biproducts. Then the following commutes for any morphisms $f_1, g_1, f_2,
g_2$ in $C$:

>>>>

\begin{definition}[Biproduct]
Suppose $\cat{C}$ a $\CMon$-enriched category. For any objects $A, B$, the \emph{biproduct} of $A$ and $B$ is
an object $A \biprod B$ of $C$ together with morphisms

@@

We say a $\CMon$-enriched $\cat{C}$ is \emph{semi-additive} if it also has finite biproducts; having products
or coproducts is sufficient.

\begin{proposition}
\label{prop:biproduct:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{itemize}
\item If $\cat{C}$ has biproducts $(A \biprod B, \biinj_A, \biinj_B, \biproj_A, \biproj_B)$ then $(A
\biprod B, \biproj_A, \biproj_B)$ is a product and $(A \biprod B, \biinj_A, \biinj_B)$ is a coproduct.
\item If $\cat{C}$ has binary products $(A \times B, \proj_1, \proj_2)$ then:
\begin{enumerate}
\item $\prodM{0_{A,B}}{0_{A,C}} = 0_{A,B \times C}$
\item $\prodM{f_1 + g_1}{f_2 + g_2} = \prodM{f_1}{f_2} + \prodM{g_1}{g_2}$
\item $(A \times B, \biinj_A, \biinj_B, \proj_1, \proj_2)$ is a
biproduct with $\biinj_A: A \to A \times B = \prodM{\id_A}{0_{A,B}}$ and $\biinj_B: B \to A \times B =
\prodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $\cat{C}$ has binary coproducts $(A + B, \inj_1, \inj_2)$ then
\begin{enumerate}
\item $\coprodM{0_{A,C}}{0_{B,C}} = 0_{A + B,C}$
\item $\coprodM{f_1 + g_1}{f_2 + g_2} = \coprodM{f_1}{f_2} + \coprodM{g_1}{g_2}$
\item $(A + B, \inj_1, \inj_2, \biproj_A, \biproj_B)$ is a
biproduct with $\biproj_A: A + B \to A = \coprodM{\id_A}{0_{B,A}}$ and $\biproj_B: A + B \to B =
\coprodM{0_{A,B}}{\id_B}$.
\end{enumerate}
\item If $\cat{C}$ has a terminal (or initial) object then it is a zero object.
\end{itemize}
\end{proposition}

\begin{proposition}
\label{prop:biproduct:prod-coprod}
Suppose $\cat{C}$ a category with biproducts. Then the following diagram in $\cat{C}$ commutes:

### 2025-04-15 appendix/notes/fam.tex
Move notes to appendix.
mathcal for metavariables that range over categories.

\subsection{Category of $I$-indexed families}

\begin{definition}[Category of $I$-indexed families of objects]
For any set $I$ and any category $C$ write $\Fam(I,C)$ for the category where:
\begin{itemize}
\item objects are the $I$-indexed families of objects of $C$;
\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $C$ for any $i \in I$.
\end{itemize}
\end{definition}

\noindent Equivalently $\Fam(I,C)$ is the functor category $\Func{I}{C}$, interpreting $I$ as a discrete
category.

\begin{definition}[Constant family]
For any object $X$ of $C$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
\end{definition}

\begin{definition}[Reindexing]
For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$ sends any
$J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
similarly for morphisms.
\end{definition}

\begin{definition}[$\Fam(-,C)$ functor]
$\Fam(-,C): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,C)$ and any function
$f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,C) \to \Fam(I,C)$.
\end{definition}

\begin{proposition}
If $C$ has limits (resp.~colimits) then $\Fam(I,C)$ has limits (resp.~colimits) computed pointwise.
\end{proposition}

>>>>

\subsection{Category of $I$-indexed families}
\label{sec:fam}

\begin{definition}[Category of $I$-indexed families of objects]
For any set $I$ and any category $\cat{C}$ write $\Fam(I,\cat{C})$ for the category where:
\begin{itemize}
\item objects are the $I$-indexed families of objects of $\cat{C}$;
\item morphisms from $X$ to $Y$ are families of morphisms $f_i: X_i \to Y_i$ in $\cat{C}$ for any $i \in I$.
\end{itemize}
\end{definition}

\noindent Equivalently $\Fam(I,\cat{C})$ is the functor category $\Func{I}{\cat{C}}$, interpreting $I$ as a
discrete category.

\begin{definition}[Constant family]
For any object $X$ of $\cat{C}$ write $\const_I(X)$ for the constant family $\{X\}_{i \in I}$.
\end{definition}

\begin{definition}[Reindexing]
For any $f: I \to J$ the \emph{reindexing} functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$ sends
any $J$-indexed family $X$ to the $I$-indexed family $X[f]$ where $X[f]_i = X_{f(i)}$ for any $i \in I$, and
similarly for morphisms.
\end{definition}

\begin{definition}[$\Fam(-,\cat{C})$ functor]
$\Fam(-,\cat{C}): \Set^{\op} \to \Cat$ is then the functor which sends any set $I$ to $\Fam(I,\cat{C})$ and
any function $f: I \to J$ to the functor $\reindex{-}{f}: \Fam(J,\cat{C}) \to \Fam(I,\cat{C})$.
\end{definition}

\begin{proposition}
If $\cat{C}$ has limits (resp.~colimits) then $\Fam(I,\cat{C})$ has limits (resp.~colimits) computed
pointwise.
\end{proposition}

### 2025-04-15 appendix/notes/set-indexed-product.tex
mathcal for metavariables that range over categories.

$C$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of $C$,
there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms $\eval_i$ in
$C$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$ in $C$, there
exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i \in I$:

@@

Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,C)$.

>>>>

$\cat{C}$ \emph{has set-indexed products} if, for any set $I$ and any $I$-indexed family $X$ of objects of
$\cat{C}$, there exists an object $\prod_{i \in I}X_i$ (also written $\prod_{I} X$) and family of morphisms
$\eval_i$ in $\cat{C}$ such that for any object $Y$ and any family of morphisms $\{f_i: Y \to X_i\}_{i \in I}$
in $\cat{C}$, there exists a unique morphism $\lambda(f)$ making the following diagram commute for every $i
\in I$:

@@

Note that $f$ is a morphism $\const_{}(Y) \to X$ in $\Fam(I,\cat{C})$.
